Deterministic Scheduling of Periodic Messages for Cloud RAN

### 2017-07-26 infocom/infocom.tex
merge

Next generations of mobile network architectures evolve toward centralized radio network architectures called C-RAN, or Cloud Radio Access Network, to reduce energy consumption costs~\cite{mobile2011c} and more generally the total cost of ownership. The main challenge for this type of architecture is to reach a latency compatible with transported protocols. The latency is measured between the sending of a message by the Remote Radio Heads (RRHs) on the field and the receptions of the answers, computed by real-time virtualized network functions of BaseBand Units (BBUs), in the cloud. For example, LTE standards requires meeting time constraints for functions like HARQ (Hybrid Automatic Repeat reQuest) that needs to be processed in $3$ms~\cite{bouguen2012lte}. In 5G, some services need end to end latency as low as 1ms~\cite{3gpp5g}. The specificity of the C-RAN context is not only the latency constraint, but also the periodicity of the data transfer in the \emph{frontaul} network between RRHs and BBUs: frames need to be emitted and received every milliseconds~\cite{bouguen2012lte}.
Our aim is to operate a C-RAN on a low-cost shared switched network.
The question we address is the following : is it possible to schedule messages emission such that there is no collision, then no buffering nor latency corresponding to queuing delays ?

Let us expose briefly our model: the network topology is modeled by a directed weighted graph and a set of paths (routes) from source nodes (RRHs) to target nodes (BBUs). Time is discretized and a unit of time or slot corresponds to the transmission of a minimal unit of data over the network. The process must be  predictable (i.e. the date of arrival of packets is known beforehand).Following LTE standard~\cite{bouguen2012lte}, in this article we assume that the process is periodic. That is, if we look at the network at times $t$ and $t+P$ where $P$ is the period, all messages are at the same position in the network. As we want to avoid any  buffering in internal nodes of the graph. Hence we need to design a periodic process with no collision between messages. To ensure that this property is always true we propose a deterministic process.  We have two sets of values that we can choose when building a periodic sending process, called a \emph{periodic assignment} in this article: the time at which each packet is sent by a RRH in each period and the waiting time in the BBU before the answer is sent back to the RRH. When building a periodic assignment, we must take into account the periodicity which makes many scheduling methods inapplicable. Not only a message must not collide with the other messages sent by the others BBU/RRH in the same period, but also in the previous and following periods.The latency, that is the time between the emission of a message and the complete return of its answer, must be minimized: this lefts budget for other latency causes, and leads to larger fiber lengths, then wider deployment areas.This means that the only buffering we are allowed -- the waiting time before sending back the answer-- must be small, in particular when the route is long. Note that the
model is technology agnostic, i.e. it is compatible with an optical network with a fixed packet size.

\footnote{Others terminologies and functional splits exist in the literature. We choose the most classical one for sake of clarity. The results of this work is fully compatible with any C-RAN architecture definition.}

Our main contributions are the following.
 In Sec.~\ref{sec:def} we propose a model of the network and the periodic sending of messages along its routes  and we introduce the star routed network that we study in this article. 
 We formalize the problem of finding a periodic assignment for sending messages without buffering (\pazl) or with buffering in the processing unit only (\pall) and shows that they are $\NP$-hard even to approximate. In Sec.~\ref{sec:PAZL}, we propose two algorithms solving \pazl when the routes between RRHs and BBUs are small or when the load is light. Using a compact representation of an assignment, we show that \pazl is fixed parameter tractable when parametrized by the number of antenna on star routed networks. Finally in Sec.~\ref{sec:PALL}, we introduce a family of algorithms to solve \pall and our experimentation allows to select one which works extremely well even in very loaded networks. In particular we show that the deterministic communication schemes we design vastly outperform the traditional statistical multiplexing with regard to latency.

@@

 Statistical multiplexing even with a large bandwidth does not comply with the latency requirements of C-RAN. Therefore, the current solution~\cite{pizzinat2015things,tayq2017real} is to use dedicated circuits for the fronthaul. Each end-point (RRH on one side, BBU on the other side) is connected through direct fiber or full optical switches. This architecture is very expensive and hardly scales in the case of a mobile network composed of tenths of thousands of deployment sites. The deterministic approach we propose has gained some traction recently: Deterministic Networking is under standardization in IEEE 802.1 TSN group~\cite{finn-detnet-architecture-08}, as well at IETF DetNet working group~\cite{ieee802}. Several patents on concepts and mechanisms for DetNet have been already published, see for example~\cite{howe2005time,leclerc2016transmission}.

@@

    \begin{figure}
       \begin{center}
      \includegraphics[width = 0.45\textwidth]{stochastic.pdf}
       %\input{figures/stochastic}
      \end{center}
      \vspace{-1cm}
      \caption{Cumulative distribution of the margin of statistical multiplexing}
      \label{fig:sto}
     \end{figure}

     We have also computed the minimum margin needed by PMLS to find a solution but we have not represented it in Figure~\ref{fig:sto} since even the third quartile was always at $0$ for a load of $100\%$. There is around $20\%$ of the instances for which the margin is larger than $0$ for a load of $100\%$ and $2\%$ for a load of $95\%$. For the instances with a margin larger than $0$, the margin is on average $400$.

>>>>

Next generations of mobile network architectures evolve toward centralized radio network architectures called C-RAN, or Cloud Radio Access Network, to reduce energy consumption costs~\cite{mobile2011c} and more generally the total cost of ownership. The main challenge for this type of architecture is to reach a latency compatible with transport protocols. The latency is measured between the sending of a message by a Remote Radio Head (RRH) and the receptions of the answer, computed by real-time virtualized network functions of a BaseBand Unit (BBU) in the cloud. For example, LTE standards requires meeting time constraints for functions like HARQ (Hybrid Automatic Repeat reQuest) that needs to be processed in $3$ms~\cite{bouguen2012lte}. In 5G, some services need end to end latency as low as 1ms~\cite{3gpp5g}. The specificity of the C-RAN context is not only the latency constraint, but also the periodicity of the data transfer in the \emph{frontaul} network between RRHs and BBUs: frames need to be emitted and received every milliseconds~\cite{bouguen2012lte}.
Our aim is to operate a C-RAN on a low-cost shared switched network.
The question we address is the following: is it possible to schedule messages such that there is no collisions to avoid latency caused by queuing delays? 
Eliminating this source of latency leaves us with more time budget for latency coming from the length of the network, and thus allows for wider deployment areas.

Let us expose briefly our model: the network topology is modeled by a directed weighted graph and a set of paths (routes) from source nodes (RRHs) to target nodes (BBUs). Time is discretized and a unit of time or slot corresponds to the transmission of a minimal unit of data over the network. The process must be  predictable (i.e. the date of arrival of packets is known beforehand). Following LTE standard~\cite{bouguen2012lte}, in this article we assume that the process is periodic. That is, if we look at the network at times $t$ and $t+P$ where $P$ is the period, all messages are at the same position in the network. As we want to avoid any  buffering in internal nodes of the graph. Hence we need to design a periodic process with no collision between messages. To ensure that this property is always true we propose a deterministic process.  We have two sets of values that we can set when building a periodic sending process, called a \emph{periodic assignment} in this article: the time at which each packet is sent by a RRH in each period and the waiting time in the BBU before the answer is sent back to the RRH. When building a periodic assignment, we must take into account the periodicity which makes many scheduling methods unusable. Not only a message must not collide with the other messages sent by the others BBU/RRH in the same period, but also in the previous and following periods. The latency, that is the time between the emission of a message and the complete return of its answer must be minimized. This means that the only buffering we are allowed -- the waiting time before sending back the answer-- must be small, in particular when the route is long. Note that the model is technology agnostic, i.e. it is compatible with an optical network with a fixed packet size.

%\footnote{Others terminologies and functional splits exist in the literature. We choose the most classical one for sake of clarity. The results of this work is fully compatible with any C-RAN architecture definition.}

Our main contributions are the following.
 In Sec.~\ref{sec:def} we propose a model of the network and the periodic sending of messages along its routes  and we introduce the star routed network that we study in this article. 
 We formalize the problem of finding a periodic assignment for sending messages without buffering (\pazl) or with buffering in the processing unit only (\pall) and shows that they are $\NP$-hard even to approximate. In Sec.~\ref{sec:PAZL}, we propose two algorithms solving \pazl when the routes between RRHs and BBUs are small or when the load is light. Using a compact representation of an assignment, we show that \pazl is fixed parameter tractable when parametrized by the number of antennas on star routed networks. Finally in Sec.~\ref{sec:PALL}, we introduce a family of algorithms to solve \pall and our experimentations allow to select one which works extremely well even in loaded networks. In particular we show that the deterministic communication schemes we design vastly outperform the traditional statistical multiplexing with regard to latency.

@@

 Statistical multiplexing even with a large bandwidth does not comply with the latency requirements of C-RAN. Therefore, the current solution~\cite{pizzinat2015things,tayq2017real} is to use dedicated circuits for the fronthaul. Each end-point, RRH or BBU, is connected through direct fiber or full optical switches. This architecture is very expensive and hardly scales in the case of a mobile network composed of tenths of thousands of deployment sites. The deterministic approach we propose has gained some traction recently: Deterministic Networking is under standardization in IEEE 802.1 TSN group~\cite{finn-detnet-architecture-08}, and in  IETF DetNet working group~\cite{ieee802}. Several patents on concepts and mechanisms for DetNet have been already published, see for example~\cite{howe2005time,leclerc2016transmission}.

@@

    \begin{figure}
       \begin{center}
      \includegraphics[width = 0.45\textwidth]{stochastic.pdf}
       %\input{figures/stochastic}
      \end{center}
      \caption{Cumulative distribution of the margin of statistical multiplexing}
      \label{fig:sto}
      \vspace{-0.5cm}
     \end{figure}

     We have also computed the minimum margin needed by PMLS to find a solution but we have not represented it in Fig.~\ref{fig:sto} since even the third quartile was always at $0$ for a load of $100\%$. There is around $20\%$ of the instances for which the margin is larger than $0$ for a load of $100\%$ and $2\%$ for a load of $95\%$. For the instances with a margin larger than $0$, the margin is on average $400$.
